---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
mydata <- read.csv(unzip("activity.zip"))
```


## What is mean total number of steps taken per day?
```{r}
mygraph.data <- aggregate(steps ~ date, mydata, sum)
hist(mygraph.data$steps, main = "Total steps per day", xlab = "Steps per day")
#calculate mean
mean(mygraph.data$steps)
#calculate median
median(mygraph.data$steps)
```



## What is the average daily activity pattern?
```{r}
mygraph.data <- aggregate(steps ~ interval, mydata, mean)
plot(mygraph.data$interval, mygraph.data$steps, type = "l", main = "Average steps per interval", xlab = "Interval", ylab = "Number of steps")
#pick interval with most steps
mygraph.data[which.max(mygraph.data$steps),]
```



## Imputing missing values
```{r}
#1. calculate number of missing values
sum(is.na(mydata$steps))
#2,3. to fill in missing values, i'll use the mean for each interval across all days
#   which is already in mygraph.data, then use that value to fill in the missing data
mydata.filled <- mydata
mydata.filled$steps[is.na(mydata.filled$steps)] <- mygraph.data$steps[match(mydata.filled$interval, mygraph.data$interval)][which(is.na(mydata.filled$steps))]
#4. histogram
mygraph.data <- aggregate(steps ~ date, mydata.filled, sum)
hist(mygraph.data$steps, main = "Total steps per day", xlab = "Steps per day")
#calculate mean
mean(mygraph.data$steps)
#calculate median
median(mygraph.data$steps)
```
Q. Do these values differ from the estimates from the first part of the assignment? 

A. Yes, but only for median value.

Q. What is the impact of imputing missing data on the estimates of the total daily number of steps?

A. Filling in the missing values, has caused data distribution to change, thus impacting the median value. it was 10765 from the original data set, and it has changed to 10766.19. the mean value for both sets stayed the same.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
# make date column into an actual date field
mydata.filled$date <- as.Date(mydata.filled$date)
# create a weekday vector
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
# classify the date column between weekend or weekday
mydata.filled$wDay <- factor((weekdays(mydata.filled$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
# create panel plot
mygraph.data <- aggregate(steps ~ interval + wDay, mydata.filled, mean)
library(ggplot2)
ggplot(mygraph.data, aes(interval, steps)) + geom_line() + facet_grid(wDay ~ .) +
    xlab("Interval") + ylab("Number of steps") + ggtitle("Average steps per interval")
```

